\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2016-05-17 paper/ms.tex
OPTIONS stuff

\subsection{Options}
A plain-text file called OPTIONS resides in the directory from which you
execute a run.

>>>>

\subsection{Options}
A plain-text file called OPTIONS resides in the directory from which you
execute a run. Here is the OPTIONS file used in the linear regression
example.

\begin{CodeChunk}
\begin{CodeInput}
# File containing parameters for DNest4
# Put comments at the top, or at the end of the line.
5       # Number of particles
10000   # new level interval
10000   # save interval
100     # threadSteps - how many steps each thread should do independently before pooling information
0       # maximum number of levels (0 = auto)
10      # Backtracking scale length (lambda in the paper)
100     # Strength of effect to force histogram to equal push. Beta in the paper
5000    # Maximum number of saves (0 = infinite)
\end{CodeInput}
\end{CodeChunk}

The first option is the number of particles per thread, here set to one.
If you use more particles
per thread, the same amount of CPU time will be spent evolving more particles,
so each one will not be evolved as far. On most problems, one particle per
thread is fine. I would only use more particles per thread when working on
a very hard problem where it is difficult for the particles to find the
important parts of parameter space.

The new level interval controls how quickly \dnest~creates new levels. In this
example, this is set to 10,000, so a new level will be created once \dnest~has
seen 10,000 likelihood values above the current top level. This is not affected by
the number of threads. It is difficult to give a sensible default for this
quantity because it depends on the complexity of the problem and how good
the Metropolis proposals are. However, 10,000
will work for many problems (I rarely use anything less than 10,000 or greater
than 300,000). Higher values are slower, but more fail-safe.

The save interval controls how often \dnest~writes a model to the output
files: this is what is usually called ``thinning''. Saving more frequently
(i.e. a smaller save interval) is usually better, although it depends on your
problem. If your models have a lot of parameters {\tt sample.txt} can get
very big and take forever to process using {\tt showresults.py}. A default
suggestion that is usually okay is to make the save interval the same as the
new level interval.

The option ``threadSteps'' controls how frequently the separate threads pool
their information about the levels. I usually use 100 or 200. If
(threadSteps) $\times$ (number of threads) is not much smaller than
saveInterval or newLevelInterval, the latter won't be what you expect, because
\dnest~only creates levels or saves particles when the threads are pooling
information.
